---
title: "Visualization alternatives"
output: html_document
date: "2023-05-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

In the original manuscript, we discussed the visualization. Here I provide some alternatives for the figures. Conclusion first: since the heterogeneity of the raw data is very high, maybe panel is still the best way, but I can add some auxiliary lines to the figure, so that readers can identify the key time point of the sub-figures. 

## Change of annual indexes

Raw data in same figure, one color for one city: 

```{r}
# idx raw
record.yr %>% 
  pivot_longer(
    cols = c(obs, user_pop, prop_long_user,day_per_user, obs_per_day), 
    names_to = "idx", 
    values_to = "idx_val"
  ) %>% 
  mutate(idx = case_when(
    idx == "obs" ~ "Observation", 
    idx == "user_pop" ~ "Population", 
    idx == "prop_long_user" ~ "Structure", 
    idx == "day_per_user" ~ "Frequency", 
    idx == "obs_per_day" ~ "Intensity"
  )) %>% 
  mutate(idx = factor(
    idx, levels = c("Observation", "Population", "Structure", "Frequency", "Intensity")
  )) %>% 
  ggplot() + 
  geom_line(aes(yr_abbr, idx_val, group = city, col = city)) + 
  facet_wrap(.~ idx, scales = "free")
```

Since the data of Tokyo is much higher than that of the other cities, we can't see the details of the other cities. Therefore, I scale the data by transfor the max value of a city into 1. 
Scaled data in same figure, one color for one city: 

```{r cars}
# idx std
record.yr %>% 
  pivot_longer(
    cols = c(obs, user_pop, prop_long_user,day_per_user, obs_per_day), 
    names_to = "idx", 
    values_to = "idx_val"
  ) %>% 
  group_by(city, idx) %>%
  mutate(idx_val_scale = idx_val / max(idx_val)) %>%
  mutate(idx = case_when(
    idx == "obs" ~ "Observation", 
    idx == "user_pop" ~ "Population", 
    idx == "prop_long_user" ~ "Structure", 
    idx == "day_per_user" ~ "Frequency", 
    idx == "obs_per_day" ~ "Intensity"
  )) %>% 
  mutate(idx = factor(
    idx, levels = c("Observation", "Population", "Structure", "Frequency", "Intensity")
  )) %>% 
  ggplot() + 
  geom_line(aes(yr_abbr, idx_val_scale, group = city, col = city)) + 
  facet_wrap(.~ idx, scales = "free")
```

Still, it is chaos ... so I think the original one (panle figure) is the best one. 

```{r fig.width=10}
# panel
record.yr %>%
  select(city, yr_abbr, user_pop, prop_long_user, day_per_user, obs_per_day) %>%
  pivot_longer(cols = c(user_pop, prop_long_user, day_per_user, obs_per_day),
               names_to = "idx", values_to = "idx_val") %>%
  mutate(idx = case_when(
    idx == "obs" ~ "Observation", 
    idx == "user_pop" ~ "Population", 
    idx == "prop_long_user" ~ "Structure", 
    idx == "day_per_user" ~ "Frequency", 
    idx == "obs_per_day" ~ "Intensity"
  )) %>% 
  mutate(idx = factor(
    idx, levels = c("Observation", "Population", "Structure", "Frequency", "Intensity")
  )) %>% 
  ggplot() +
  geom_line(
    aes(as.numeric(as.character(yr_abbr)), idx_val)
  ) + 
  facet_grid2(vars(idx), vars(city), scales = "free", independent = "y")
```

The problem of the panel figure is that it is difficult to identify if the cities change in a same way, for instance, if they all decrease since 2019. Therefore, I added a auxiliary line (showing year 2019) to each figure. 

```{r fig.width=10}
record.yr %>%
  select(city, yr_abbr, user_pop, prop_long_user, day_per_user, obs_per_day) %>%
  pivot_longer(cols = c(user_pop, prop_long_user, day_per_user, obs_per_day),
               names_to = "idx", values_to = "idx_val") %>%
  mutate(idx = case_when(
    idx == "obs" ~ "Observation", 
    idx == "user_pop" ~ "Population", 
    idx == "prop_long_user" ~ "Structure", 
    idx == "day_per_user" ~ "Frequency", 
    idx == "obs_per_day" ~ "Intensity"
  )) %>% 
  mutate(idx = factor(
    idx, levels = c("Observation", "Population", "Structure", "Frequency", "Intensity")
  )) %>% 
  ggplot() +
  geom_line(
    aes(as.numeric(as.character(yr_abbr)), idx_val)
  ) +
  geom_vline(xintercept = 19, col = "red", alpha = 0.5) + 
  facet_grid2(vars(idx), vars(city), scales = "free", independent = "y")
```

## LMDI result

For LMDI result, I also provide some alternatives. 

The original one: 

```{r}
lmdi %>%
    select(-o_0, -o_t, -delt_o) %>%
    pivot_longer(cols = c(delt_p, delt_s, delt_f, delt_i),
                 names_to = "delt", values_to = "delt_val") %>%
    group_by(city, year_t, year_0) %>%
    mutate(delt_abs_max = max(abs(delt_val))) %>%
    ungroup() %>%
    mutate(delt_val_scale = delt_val / delt_abs_max) %>%
    mutate(pos_neg = case_when(
      delt_val_scale > 0 ~ 1,
      TRUE ~ -1
    )) %>%
    mutate(delt = factor(delt, levels = c("delt_p", "delt_s", "delt_f", "delt_i"))) %>%
    ggplot() +
    geom_col(aes(year_t, delt_val_scale, fill = as.character(pos_neg))) +
    theme_bw() +
    facet_grid(city ~ delt) +
    scale_fill_manual(
      name = "Effect Direction",
      limits = c("1", "-1"),
      values = c("darkgreen", "darkred"),
      labels = c("Positive", "Negative")
    ) +
    theme(legend.position = "bottom") +
    scale_y_continuous(limits = c(-1, 1), breaks = seq(-1, 1, 1)) +
    labs(x = "", y = "Scaled Effect")
```

The line plot, which is usually applied in LMDI studies. We can do that either by effect: 

```{r}
lmdi %>%
    select(-o_0, -o_t, -delt_o) %>%
    pivot_longer(cols = c(delt_p, delt_s, delt_f, delt_i),
                 names_to = "delt", values_to = "delt_val") %>%
    group_by(city, year_t, year_0) %>%
    mutate(delt_abs_max = max(abs(delt_val))) %>%
    ungroup() %>%
    mutate(delt_val_scale = delt_val / delt_abs_max) %>%
    mutate(delt = factor(delt, levels = c("delt_p", "delt_s", "delt_f", "delt_i"))) %>%
    ggplot() +
    geom_line(aes(year_t, delt_val_scale, group = city, col = city)) +
    theme_bw() +
    facet_wrap(.~ delt)
```

Or by cities: 

```{r}
lmdi %>%
  select(-o_0, -o_t, -delt_o) %>%
  pivot_longer(cols = c(delt_p, delt_s, delt_f, delt_i),
               names_to = "delt", values_to = "delt_val") %>%
  group_by(city, year_t, year_0) %>%
  mutate(delt_abs_max = max(abs(delt_val))) %>%
  ungroup() %>%
  mutate(delt_val_scale = delt_val / delt_abs_max) %>%
  mutate(delt = factor(delt, levels = c("delt_p", "delt_s", "delt_f", "delt_i"))) %>%
  ggplot() +
  geom_line(aes(year_t, delt_val_scale, group = delt, col = delt)) +
  theme_bw() +
  facet_wrap(.~ city)
```

I think the best one is using tile plot: 

```{r fig.width=10}
lmdi %>%
    select(-o_0, -o_t, -delt_o) %>%
    pivot_longer(cols = c(delt_p, delt_s, delt_f, delt_i),
                 names_to = "delt", values_to = "delt_val") %>%
    group_by(city, year_t, year_0) %>%
    mutate(delt_abs_max = max(abs(delt_val))) %>%
    ungroup() %>%
    mutate(delt_val_scale = delt_val / delt_abs_max) %>%
    mutate(delt = factor(delt, levels = c("delt_p", "delt_s", "delt_f", "delt_i"))) %>%
    ggplot(aes(year_t, city)) +
    geom_tile(aes(fill = delt_val_scale)) +
    # theme_bw() +
    scale_fill_gradient2(
      name = "Effect", low = "darkred", high = "darkgreen", mid = "white"
    ) +
    geom_text(aes(label = sprintf("%.1f", delt_val_scale))) +
    facet_wrap(.~ delt, nrow = 1) +
    labs(x = "", y = "City")
```
